---
title: "Productivity Tools"
author: "natsuMishi"
date: "2023-12-31"
output: html_document
---

# Section 1 Overview

The Installing Software section walks you through the steps to download and install R, RStudio, git (and git bash on Windows machines), create a GitHub account, and connect RStudio to GitHub.

### Installing Software 

-   R: the programming language we use to analyze data

-   RStudio: the integrated desktop environment we use to edit, organize, and test R scripts

-   Git (and Git Bash for Windows): version control system

### Installing R 

-   You need to install R before using RStudio, which is an interactive desktop environment.

-   Select base subdirectory in CRAN and click download.

-   Select all default choices in the installation process.

-   We recommend selecting English for language to help you better follow the course.

-   One can try using the R console, but for productivity purposes, we can switch to RStudio.

### Installing RStudio 

-   You can download the latest version of RStudio at the [RStudio website External link](https://www.rstudio.com/products/rstudio/download/).

-   The free desktop version is more than enough for this course.

-   Make sure to choose the version for your own operating system.

-   Choose "Yes" for all defaults in the installation process.

### Using RStudio for the first time 

-   In Windows, search for Rstudio in "Start"

-   In Mac, choose Rstudio from applications, or use Command+Space to use spotlight search to avoid using the mouse

-   To start a new script: file \> new file \> Rscript (or using key bindings: Ctrl+Shift+N on Windows, and Command+Shift+N on Mac)

-   Key bindings from Rstudio cheatsheet can be located with Help \> cheatsheet \> Rstudio IDE cheatsheet

### Installing packages 

-   We will be using **tidyverse** and **dslabs** packages for this course, which do not come preinstalled in base R.

-   Install packages from R console: `install.packages("pkg_name")`

-   Install packages from Rstudio interface: Tools \> install packages (allow autocomplete)

-   Once installed, we can use `library("pkg_name")` to load a package each time we want to use it

-   Check the packages you have installed

-   Good practice: make a script to install all the packages you need for a fresh R

#### Code

```{r}
install.packages("dslabs")  # to install a single package
install.packages(c("tidyverse", "dslabs")） # to install two packages at the same time
installed.packages() # to see the list of all installed packages
```

### Running commands while editing scripts 

-   RStudio has many useful features as an R editor, including the ability to test code easily as we write scripts and several autocomplete features.

-   Keyboard shortcuts:

    -   Save a script: Ctrl+S on Windows and Command+S on Mac

    -   Run an entire script:  Ctrl+Shift+Enter on Windows Command+Shift+Return on Mac, or click "Source" on the editor pane

    -   Run a single line of script: Ctrl+Enter on Windows and Command+Return on Mac while the cursor is pointing to that line, or select the chunk and click "run"

    -   Open a new script: Ctrl+Shift+N on Windows and Command+Shift+N on Mac

#### Code

```{r}
library(tidyverse)
```

### Keeping Organized with RStudio Projects 

-   RStudio provides a way to keep all the components of a data analysis project organized into one folder and to keep track of information about this project.

-   To start a project, click on File \> New Project \> New repository \> New project \> decide the location of files and give a name to the project, e.g. "my-first-project". This will then generate a Rproj file called my-first-project.Rproj in the folder associated with the project, from which you can double click to start where you last left off.

-   The project name will appear in the upper left corner or the upper right corner, depending on your operating system. When you start an RStudio session with no project, it will display "Project: (None)".

### Installing Git Introduction 

-   Git is a version control system, tracking changes and coordinating the editing of code.

-   GitHub is a hosting system for code, which can help with your career profile.

-   Git is most effectively used with Unix, but it can also interface with RStudio.

### Install on Windows

1.  Download Git bash from [https://git-scm.com/ External link](https://git-scm.com/)

2.   When asked to choose the default editor for Git, we recommend choosing nano if you do not already know VIM.

3.   The "git and optional Unix tools from Windows" option will allow you to learn Unix from Rstudio, however, it might interfere with the Windows command line.

4.   Finally, change the Rstudio preference so that you are using Git bash as the terminal (only for Windows user).

### GitHub 

-   Sign up for a GitHub account, with a name that is professional, short, and easy to remember

-   Connect to RStudio: global options \> Git/SVM, enter the path to git executables

-   To avoid typing our GitHub password every time, we create a SSH/RSA key automatically through RStudio with the *create RSA key* button.

### GitHub Repositories

To create a new GitHub repository and link it to an RStudio project, follow 2 steps:

Step 1: Initialize a new repo on GitHub by clicking repository \> new \> choose a descriptive name.

Step 2: Connect to RStudio (next video).
