---
title: '1.4 Assessment: Discrete Probability'
author: "natsuMishi"
date: "2023-12-24"
output: html_document
---

**Questions 3 and 4: Esophageal cancer and alcohol/tobacco use, part 1**

Case-control studies help determine whether certain exposures are associated with outcomes such as developing cancer. The built-in dataset esoph contains data from a case-control study in France comparing people with esophageal cancer (cases, counted in ncases) to people without esophageal cancer (controls, counted in ncontrols) that are carefully matched on a variety of demographic and medical characteristics. The study compares alcohol intake in grams per day (alcgp) and tobacco intake in grams per day (tobgp) across cases and controls grouped by age range (agegp).

The dataset is available in base R and can be called with the variable name esoph:

```{r}
options(digits = 3)
library(gtools)
library(tidyverse)
library(dplyr)
head(esoph)
```

You will be using this dataset to answer the following four multi-part questions (Questions 3-6).

The following three parts have you explore some basic characteristics of the dataset.

Each row contains one group of the experiment. Each group has a different combination of age, alcohol consumption, and tobacco consumption. The number of cancer cases and number of controls (individuals without cancer) are reported for each group.

**Question 3a**

**How many groups are in the study?**

Since each row contains one group of the experiment:

```{r}
nrow(esoph)
```

**Question 3b**

**How many cases are there?**

```{r}
all_cases <- sum(esoph$ncases)
all_cases
```

**Question 3c**

**How many controls are there?**

```{r}
all_controls <- sum(esoph$ncontrols)
all_controls
```

The following four parts ask you to explore some probabilities within this dataset related to alcohol and tobacco consumption.

**Question 4a**

What is the probability that a subject in the highest alcohol consumption group is a cancer case?

The highest alcohol consumption group is 120+

```{r}
#To find the highest alcohol consumption group
levels(esoph$alcgp)

#To find the count of cancer cases in this group
count_cancer_and_120 <- esoph %>%
  filter(esoph$alcgp == "120+") %>%
  .$ncases %>%
  sum()

#To find the count of control cases in this group
count_control_and_120 <- esoph %>%
  filter(esoph$alcgp == "120+") %>%
  .$ncontrol %>%
  sum()

count_cancer_and_120
count_control_and_120

#To calculate the probability that a subject in the highest alcohol consumption group is a cancer case
prob_cancer_and_120 <- count_cancer_and_120/(count_cancer_and_120+count_control_and_120)

prob_cancer_and_120
```

**Question 4b**

What is the probability that a subject in the lowest alcohol consumption group is a cancer case?

```{r}
#To find the lowest alcohol consumption group
levels(esoph$alcgp)

#To find the count of cancer cases in this group
count_cancer_and_lowest <- esoph %>%
  filter(esoph$alcgp == "0-39g/day") %>%
  .$ncases %>%
  sum()

#To find the count of control cases in this group
count_control_and_lowest <- esoph %>%
  filter(esoph$alcgp == "0-39g/day") %>%
  .$ncontrol %>%
  sum()

count_cancer_and_lowest
count_control_and_lowest

#To calculate the probability that a subject in the highest alcohol consumption group is a cancer case
prob_cancer_and_lowest <- count_cancer_and_lowest/(count_cancer_and_lowest+count_control_and_lowest)

prob_cancer_and_lowest
```

**Question 4c**

Given that a person is a case, what is the probability that they smoke 10g or more a day?

```{r}
#Probability of cancer
count_cancer <- esoph %>% summarize(count_cancer = sum(ncases))
count_cancer

#Probability of cancer and 10g or more
count_cancer_10ormore <- esoph %>% filter(!tobgp %in% c("0-9g/day")) %>%
  summarize(smoke10ormore_cancer = sum(ncases))
count_cancer_10ormore

prob_10ormore_given_cancer <- count_cancer_10ormore/count_cancer
prob_10ormore_given_cancer
```

**Question 4d**

Given that a person is a control, what is the probability that they smoke 10g or more a day?

```{r}
#Probability of control
count_controls <- esoph %>% summarize(count_controls = sum(ncontrols))
count_controls

#Probability of cancer and 10g or more
count_controls_10ormore <- esoph %>% filter(!tobgp %in% c("0-9g/day")) %>%
  summarize(smoke10ormore_controls = sum(ncontrols))
count_controls_10ormore

prob_10ormore_given_control <- count_controls_10ormore/count_controls
prob_10ormore_given_control
```

**Question 5a**

For cases, what is the probability of being in the highest alcohol group?

```{r}
#Probability of cancer
count_cases <- esoph %>% summarize(count_cases = sum(ncases))
count_cases

#Probability of cancer and 10g or more
count_cases_120 <- esoph %>% filter(alcgp == "120+") %>%
  summarize(count_cases_120 = sum(ncases))
count_cases_120

count_cases_120/count_cases
```

**Question 5b**

For cases, what is the probability of being in the highest tobacco group?

```{r}
#Probability of cancer
count_cases <- esoph %>% summarize(count_cases = sum(ncases))
count_cases

#Probability of cancer highest tobbaco group
count_cases_30 <- esoph %>% filter(tobgp == "30+") %>%
  summarize(count_cases_30 = sum(ncases))
count_cases_30

count_cases_30/count_cases
```

**Question 5c**

For cases, what is the probability of being in the highest alcohol group and the highest tobacco group?

```{r}
esoph %>% filter(alcgp == "120+" & tobgp == "30+") %>%
  summarize(count = sum(ncases))
all_cases <- esoph %>% summarize(count = sum(ncases))

10/all_cases
```

**Question 5d**

For cases, what is the probability of being in the highest alcohol group or the highest tobacco group?

```{r}
esoph %>% filter(alcgp == "120+" | tobgp == "30+") %>%
  summarize(count = sum(ncases))
all_cases <- esoph %>% summarize(count = sum(ncases))

66/all_cases
```

**Question 6a**

For controls, what is the probability of being in the highest alcohol group?

```{r}
esoph %>% filter(alcgp == "120+") %>%
  summarize(count = sum(ncontrols))
all_controls <- esoph %>% summarize(count = sum(ncontrols))
22/all_controls
```

**Question 6b**

How many times more likely are cases than controls to be in the highest alcohol group?

```{r}
esoph %>% filter(alcgp == "120+") %>%
  summarize(
    prob_ca = sum(ncases)/200,
    prob_co = sum(ncontrols)/775,
    ratio = prob_ca/prob_co
  )

```

**Question 6c**

For controls, what is the probability of being in the highest tobacco group?

```{r}
esoph %>% filter(tobgp == "30+") %>%
  summarize(sum(ncontrols))
all_controls
51/775
```

**Question 6d**

For controls, what is the probability of being in the highest alcohol group and the highest tobacco group?

```{r}
esoph %>% filter(tobgp == "30+" & alcgp == "120+") %>%
  summarize(sum(ncontrols))
3/all_controls
```

**Question 6e**

For controls, what is the probability of being in the highest alcohol group or the highest tobacco group?

```{r}
esoph %>% filter(tobgp == "30+" | alcgp == "120+") %>%
  summarize(sum(ncontrols))
70/all_controls
```

**Question 6f**

How many times more likely are cases than controls to be in the highest alcohol group or the highest tobacco group?

```{r}
esoph %>% filter(alcgp == "120+" | tobgp == "30+") %>%
  summarize(
    prob_co = sum(ncontrols)/775,
    prob_ca = sum(ncases)/200,
    prob_ca/prob_co
  )

all_controls
all_cases

```
