---
title: "Brexit Poll Analysis"
author: "natsuMishi"
date: "2024-01-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Brexit poll analysis - Part 1

### Overview

In June 2016, the United Kingdom (UK) held a referendum to determine whether the country would “Remain” in the European Union (EU) or “Leave” the EU. This referendum is commonly known as Brexit. Although the media and others interpreted poll results as forecasting “Remain” ( 𝑝\>0.5) , the actual proportion that voted “Remain” was only 48.1% (𝑝=0.481) and the UK thus voted to leave the EU. Pollsters in the UK were criticized for overestimating support for “Remain”.

### Important definitions

#### Data Import

Import the `brexit_polls` polling data from the **dslabs** package and set options for the analysis:

```{r}
# suggested libraries and options
library(tidyverse)
options(digits = 3)

# load brexit_polls object
library(dslabs)
data(brexit_polls)
```

#### Final Brexit parameters

Define 𝑝=0.481 as the actual percent voting “Remain” on the Brexit referendum and 𝑑=2𝑝−1=−0.038 as the actual spread of the Brexit referendum with “Remain” defined as the positive outcome:

```{r}
p <- 0.481    # official proportion voting "Remain"
d <- 2*p-1    # official spread
```

### Question 1: Expected value and standard error of a poll

The final proportion of voters choosing “Remain” was 𝑝=0.481 . Consider a poll with a sample of 𝑁=1500 voters.

What is the expected total number of voters in the sample choosing “Remain”?

```{r}
N <- 1500
N*p
```

What is the standard error of the total number of voters in the sample choosing “Remain”?

```{r}
sqrt(N*p*(1-p))
```

What is the expected value of 𝑋̂ , the proportion of “Remain” voters?

```{r}
p
```

What is the standard error of 𝑋̂ , the proportion of “Remain” voters?

```{r}
sqrt(p*(1-p)/N)
```

What is the expected value of 𝑑 , the spread between the proportion of “Remain” voters and “Leave” voters?

```{r}
2*p-1
```

What is the standard error of 𝑑 , the spread between the proportion of “Remain” voters and “Leave” voters?

```{r}
2*sqrt(p*(1-p)/N)
```

### Question 2: Actual Brexit poll estimates

Load and inspect the brexit_polls dataset from dslabs, which contains actual polling data for the 6 months before the Brexit vote. Raw proportions of voters preferring “Remain”, “Leave”, and “Undecided” are available (remain, leave, undecided) The spread is also available (spread), which is the difference in the raw proportion of voters choosing “Remain” and the raw proportion choosing “Leave”.

Calculate x_hat for each poll, the estimate of the proportion of voters choosing “Remain” on the referendum day ( 𝑝=0.481 ), given the observed spread and the relationship 𝑑̂ =2𝑋̂ −1 . Use mutate() to add a variable x_hat to the brexit_polls object by filling in the skeleton code below:

What is the average of the observed spreads (spread)?

```{r}
mean(brexit_polls$spread)
```

What is the standard deviation of the observed spreads?

```{r}
sd(brexit_polls$spread)
```

What is the average of x_hat, the estimates of the parameter

```{r}
brexit_polls <- brexit_polls %>% 
                  mutate(x_hat = (spread + 1) /2)
mean(brexit_polls$x_hat)
```

What is the standard deviation of x_hat?

```{r}
sd(brexit_polls$x_hat)
```

### Question 3: Confidence interval of a Brexit poll

Consider the first poll in brexit_polls, a YouGov poll run on the same day as the Brexit referendum:

Use qnorm() to compute the 95% confidence interval for 𝑋̂.

What is the lower bound of the 95% confidence interval?

```{r}
x_hat2 <- brexit_polls[1,]$x_hat
N <- brexit_polls[1,]$samplesize
se2 <- sqrt(x_hat2 * (1 - x_hat2) / N)
x_hat2 - qnorm(0.975) * se2
```

What is the upper bound of the 95% confidence interval?

```{r}
x_hat2 + qnorm(0.975) * se2
```

Does the 95% confidence interval predict a winner (does not cover p=0.5)? Does the 95% confidence interval cover the true value of p observed during the referendum?

```{r}
!between(0.5, x_hat2 - qnorm(.975)*se2, x_hat2 + qnorm(.975)*se2)    # predicts winner
between(0.481, x_hat2 - qnorm(.975)*se2, x_hat2 + qnorm(.975)*se2)    # does not cover p
```

# Brexit poll analysis - Part 2

```{r}
# suggested libraries
library(tidyverse)

# load brexit_polls object and add x_hat column
library(dslabs)
data(brexit_polls)
brexit_polls <- brexit_polls %>%
    mutate(x_hat = (spread + 1)/2)

# final proportion voting "Remain"
p <- 0.481
```

### Question 4: Confidence intervals for polls in June

Create the data frame june_polls containing only Brexit polls ending in June 2016 (enddate of “2016-06-01” and later). We will calculate confidence intervals for all polls and determine how many cover the true value of 𝑑.

First, use mutate() to calculate a plug-in estimate se_x_hat for the standard error of the estimate SE^[1](https://rstudio-pubs-static.s3.amazonaws.com/709447_61b5af1beb004f55abbb7e6927b66a04.html#fn1)^ for each poll given its sample size and value of 𝑋̂ (x_hat). Second, use mutate() to calculate an estimate for the standard error of the spread for each poll given the value of se_x_hat. Then, use mutate() to calculate upper and lower bounds for 95% confidence intervals of the spread. Last, add a column hit that indicates whether the confidence interval for each poll covers the correct spread 𝑑=−0.038.

How many polls are in june_polls?

```{r}
june_polls <- brexit_polls %>% filter(enddate >= '2016-06-01') 
nrow(june_polls)
```

What proportion of polls have a confidence interval that covers the value 0?

```{r}
d <- -0.38
june_polls <- june_polls %>%
  mutate(se_x_hat = sqrt(x_hat * (1-x_hat)/ samplesize), se_spread = 2*se_x_hat,
         lower = spread - qnorm(0.975) * se_spread,
         upper = spread + qnorm(0.975) * se_spread,
         hit = (lower < 2*p-1 & upper > 2*p-1))
mean(june_polls$lower < 0 & june_polls$upper > 0)
```

# Brexit poll analysis - Part 2

```{r}
# suggested libraries
library(tidyverse)

# load brexit_polls object and add x_hat column
library(dslabs)
data(brexit_polls)
brexit_polls <- brexit_polls %>%
    mutate(x_hat = (spread + 1)/2)

# final proportion voting "Remain"
p <- 0.481
```

### Question 4: Confidence intervals for polls in June 

```{r}
d <- -0.038 #correct spread
june_polls <- brexit_polls %>%
  filter(enddate >= "2016-06-01") %>%
  mutate(se_x_hat = sqrt(x_hat * (1-x_hat)/samplesize)) %>%
  mutate(se_spread = 2*se_x_hat) %>%
  mutate(
    lower = spread - qnorm(0.975)*se_spread,
    upper = spread + qnorm(0.975)*se_spread,
    hit = (lower < d & upper > d)
  )

#How many polls are in june_polls?
length(june_polls$pollster)

#What proportion of polls have a confidence interval that covers the value 0?
covers_0 <- june_polls %>% mutate(cover_0 = lower < 0 & upper > 0) %>% .$cover_0
class(covers_0)
mean(covers_0)

#What proportion of polls predict "Remain" (confidence interval entirely above 0?)
remain <- june_polls %>% mutate(remain = lower > 0 & upper > 0) %>% .$remain
mean(remain)

#What proportions of polls have a confidence interval covering the true value of d?
mean(june_polls$hit)
```

### Question 5: Hit rate by pollster

Group and summarize the `june_polls` object by pollster to find the proportion of hits for each pollster and the number of polls per pollster. Use `arrange()` to sort by hit rate.

```{r}
june_polls %>% 
  group_by(pollster) %>%
  summarize(proportion_hits = mean(hit == TRUE), polls = n()) %>% arrange(desc(proportion_hits))
  
```

### Question 6: Boxplot of Brexit polls by poll type 

Make a boxplot of the spread in `june_polls` by poll type.

```{r}
june_polls %>% ggplot(aes(poll_type, spread)) + 
  geom_boxplot() +
  geom_jitter()
```

### Question 7: Combined spread across poll type 

Calculate the confidence intervals of the spread combined across all polls in `june_polls`, grouping by poll type. Recall that to determine the standard error of the spread, you will need to double the standard error of the estimate.

Use this code (which determines the total sample size per poll type, gives each spread estimate a weight based on the poll's sample size, and adds an estimate of p from the combined spread) to begin your analysis:

```{r}
combined_by_type <- june_polls %>%
        group_by(poll_type) %>%
        summarize(N = sum(samplesize),
                  spread = sum(spread*samplesize)/N,
                  p_hat = (spread + 1)/2)

#What is the lower bound of the 95% confidence interval for online voters?
spread_online <- combined_by_type$spread[1]

se_p_hat_online <- sqrt(p_hat_online * (1-p_hat_online)/combined_by_type$N[1])

spread_online - qnorm(0.975)* 2* se_p_hat_online 

#What is the upper bound of the 95% confidence interval for online voters?
spread_online + qnorm(0.975)* 2 * se_p_hat_online 


#Code from edX:
combined_by_type <- june_polls %>%
    group_by(poll_type) %>%
    summarize(N = sum(samplesize),
              spread = sum(spread*samplesize)/N,
              p_hat = (spread + 1)/2,
              se_spread = 2*sqrt(p_hat*(1-p_hat)/N),
              spread_lower = spread - qnorm(.975)*se_spread,
              spread_upper = spread + qnorm(.975)*se_spread)

combined_by_type %>%
    filter(poll_type == "Online") %>%
    pull(spread_lower)

combined_by_type %>%
    filter(poll_type == "Online") %>%
    pull(spread_upper)
```

### Question 8: Interpreting combined spread estimates across poll type 

Interpret the confidence intervals for the combined spreads for each poll type calculated in the previous problem.

Which of the following are TRUE about the confidence intervals of the combined spreads for different poll types?

```{r}
combined_by_type <- june_polls %>%
        group_by(poll_type) %>%
        summarize(N = sum(samplesize),
                  spread = sum(spread*samplesize)/N,
                  p_hat = (spread + 1)/2)
combined_by_type %>%
  mutate(lower = spread - qnorm(.975)*2*sqrt(p_hat*(1-p_hat)/N),
         upper = spread + qnorm(.975)*2*sqrt(p_hat*(1-p_hat)/N),
         ci_amplitude = abs(upper-lower),
         hit_true_d = lower <= -0.038 & upper >= -0.038
         )
```

# Brexit poll analysis - Part 3

```{r}
# suggested libraries
library(tidyverse)

# load brexit_polls object and add x_hat column
library(dslabs)
data(brexit_polls)
brexit_polls <- brexit_polls %>%
    mutate(x_hat = (spread + 1)/2)

# final proportion voting "Remain"
p <- 0.481
```

### Question 9: Chi-squared p-value 

```{r}
# This code computes the confidence intervals for all Brexit polls in 2016 and then calculates whether the confidence intervals covers the actual value of the spread d = -0.038.
brexit_hit <- brexit_polls %>%
  mutate(p_hat = (spread + 1)/2,
         se_spread = 2*sqrt(p_hat*(1-p_hat)/samplesize),
         spread_lower = spread - qnorm(.975)*se_spread,
         spread_upper = spread + qnorm(.975)*se_spread,
         hit = spread_lower < -0.038 & spread_upper > -0.038) %>%
  select(poll_type, hit)

#Use brexit_hit to make a two-by-two table of poll type and hit status. Then use the chisq.test() function to perform a chi-squared test to determine whether the difference in hit rate is significant. 
two_by_two <- t(table(brexit_hit$poll_type, brexit_hit$hit))
two_by_two

#What is the p-value of the chi-squared test comparing the hit rate of online and telephone polls?
chisq.test(two_by_two)

#edX code
brexit_chisq <- table(brexit_hit$poll_type, brexit_hit$hit)
chisq.test(brexit_chisq)$p.value
 # online > telephone
hit_rate <- brexit_hit %>%
    group_by(poll_type) %>%
    summarize(avg = mean(hit))
hit_rate$avg[hit_rate$poll_type == "Online"] > hit_rate$avg[hit_rate$poll_type == "Telephone"]

# statistically significant
chisq.test(brexit_chisq)$p.value < 0.05
```

### Question 10: Odds ratio of online and telephone poll hit rate 

Use the two-by-two table constructed in the previous exercise to calculate the odds ratio between the hit rate of online and telephone polls to determine the magnitude of the difference in performance between the poll types.

Calculate the odds that an online poll generates a confidence interval that covers the actual value of the spread.

```{r}
two_by_two
odds_online_true <- (48/(37+48)) / (37/(37+48))
odds_online_true
```

Calculate the odds that a telephone poll generates a confidence interval that covers the actual value of the spread.

```{r}
odds_telephone_true <- (10/(32+10))/(32/(32+10))
odds_telephone_true
```

Calculate the odds ratio to determine how many times larger the odds are for online polls to hit versus telephone polls.

```{r}
odds_ratio <- (32*48)/(37*10)
odds_ratio
```

### Question 11: Plotting spread over time 

### Question 12: Plotting raw percentages over time 

Use the following code to create the object `brexit_long`, which has a column `vote` containing the three possible votes on a Brexit poll ("remain", "leave", "undecided") and a column `proportion` containing the raw proportion choosing that vote option on the given poll:

```{r}
brexit_long <- brexit_polls %>%
    gather(vote, proportion, "remain":"undecided") %>%
    mutate(vote = factor(vote))
```

Make a graph of proportion over time colored by vote. Add a smooth trendline with `geom_smooth()` and `method = "loess"` with a span of 0.3.

Which of the following are TRUE?

```{r}
brexit_polls %>% ggplot(aes(enddate, spread, color = poll_type)) +
  geom_smooth(method = "loess", span = 0.4) + 
  geom_point() +
  geom_hline(yintercept = d)

brexit_polls
```

### Question 12: Plotting raw percentages over time 

Use the following code to create the object `brexit_long`, which has a column `vote` containing the three possible votes on a Brexit poll ("remain", "leave", "undecided") and a column `proportion` containing the raw proportion choosing that vote option on the given poll:

```{r}
brexit_long <- brexit_polls %>%
    gather(vote, proportion, "remain":"undecided") %>%
    mutate(vote = factor(vote))
```

Make a graph of proportion over time colored by vote. Add a smooth trendline with `geom_smooth()` and `method = "loess"` with a span of 0.3.

Which of the following are TRUE?\

```{r}
brexit_long %>% ggplot(aes(enddate, proportion, color = vote)) + 
  geom_smooth(method = "loess", span = 0.3)

```
